### 2011-10-09 doc/theses/vitasek/chapter-1-definitions.tex
Thesis: in progress

TODO we shall introduce a number of definitions, some only auxiliary, some necessary in the whole work

@@

@@

\begin{define}[Maximum weighted independent set]
Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied.
\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v in I_{max}} w(v)\]
\end{define}

TODO finding this is NP-hard - cite

@@

TODO define LP rigorously

@@

TODO it is possible to formulate minimization function

>>>>

@@

\begin{define}[$name()$]
	Given an attribute mapping $m = M_{x}^{y}$, $name(m)$ shall be defined as the string $x-y$.
\end{define}

@@

\begin{define}[Maximum weighted independent set]
Given an undirected graph $G = (V, E)$ and a weight function $w: V \rightarrow \mathbb{R}$, an independent set $I_{max}$ is the \textit{maximum weighted independent set}, iff the following is satisfied.
\[\forall I' \subseteq V, I' \text{is an independent set}: \sum_{v \in I'} w(v) \leqslant \sum_{v \in I_{max}} w(v)\]
\end{define}

It is well known that finding the maximum weighted IS is an NP-hard optimization problem, see e.g. \cite{JM1986425}.

@@

The problem of \textit{linear programming} is optimization of a linear function under a set of linear constraints. The formulation is usually called a \textit{linear program}. It can be written in the following form (minimization version is possible too).

@@

Solving a linear program is usually possible in polynomial time using the \textit{simplex algorithm} described for example in \cite{dantzig1998linear}.

\section{Mixed Integer Problem}

\begin{define}[Mixed integer problem]
	MIP, or \textit{mixed integer problem}, is an instance of linear programming in which some or all variables are limited to integral or boolean (0, 1) values.
\end{define}

Solving MIP in general is $\NP$-hard.

### 2011-10-09 doc/theses/vitasek/chapter-3-mip.tex
Thesis: in progress

We have shown that ID set -> IS, now we show the translation to MIP formulation and solve it using GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}

TODO MIP formulation: LP is simple using Simplex method, however as soon as we require integrality or boolean values, it is NP-hard.

\begin{define}[Mixed integer problem]
	MIP, or \textit{mixed integer problem}, is an instance of linear programming in which some variables are limited to integral or boolean (0, 1) values.
\end{define}

TODO show how to translate IS to MIP

\section{Finding ID sets with GLPK}

TODO how GLPK works (branch \& bound), how to construct a problem, how output looks, how to interpret it, limit time-> limit quality, ...

\begin{define}[$name()$]
Given an attribute mapping $m = M_{x}^{y}$, $name(m)$ shall be defined as the string $x\-y$.
\end{define}

@@

TODO equations for the MIP formulation

For each AM $m$ there will be one binary variable $x_{name(m)}$. Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings. For each pair of AMs $m_1, m_2$ that cannot share the same ID set (because of conflicting type or image) there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$. $\mathbf{b}$ will be a vector of ones with corresponding length.

@@

The problem is, recall, to solve the following.

@@

In GLPK \textit{mathprog} language, it translates to the following formulation. % TODO cite mathprog

\begin{scriptsize}
\begin{verbatim}
set AMs;
param Weight {i in AMs};
var x {i in AMs} binary;
maximize z: sum {i in AMs} x[i] * Weight[i];
s.t. c1: x['y-b'] + x['y-c'] <= 1;
s.t. c2: x['y-a'] + x['y-c'] <= 1;
s.t. c3: x['y-a'] + x['y-b'] <= 1;
s.t. c4: x['y-a'] + x['z-a'] <= 1;
data;
set AMs := y-c y-b y-a z-a;
param Weight :=
y-c 0.2
y-b 0.2
y-a 0.6
z-a 0.4;
end;
\end{verbatim}
\end{scriptsize}

@@

This trivial heuristics sorts all candidate AMs by their decreasing weights (TODO link definition of weight) and then tries to iteratively add them to solution, if possible. This way it can create only one solution, and again, this is a hungry heuristic.

>>>>

In this chapter we introduce a new approach to finding maximum ID sets. First, we transform the problem formulation to maximum weighted IS problem formulation. Then we transform this into a MIP formulation, and demonstrate how this can be solved using a solver such as GLPK. \nomenclature{GLPK}{GNU Linear Programming Kit}
We will continue by applying heuristical approaches to improve the performance of the process.

\section{ID Set to IS Formulation}

Given $C = \{m_1, \ldots, m_n\}$ a set of all AMs in a document, we construct a graph $G = (V,E)$ as follows. For each AM $m_i \in C$ we create a vertex $v_{name(m_i)}$. Two vertices $v_{name(m_i)}$ and $v_{name(m_j)}$ shall be connected by an edge iff they cannot share the same ID set, either because they have the same type ($\tau(m_i) = \tau(m_j)$), or their images intersect ($\iota(m_i) \cap \iota(m_j) \neq \emptyset$). Weight of a vertex $v_{name(m_i)}$ is the weight of the attribute mapping: $w(v_{name(m_i)}) = weight(m_i)$.

Now finding the maximum weighted IS in $G$ finds the maximum (optimal) ID set in the original document.

\section{IS to MIP Formulation}

Given a graph $G = (V,E)$ with a weight function $w: V \rightarrow \mathbb{R}$, we introduce a binary variable $x_i$ for each vertex $v_i \in V$ and an inequality constraint $x_i + x_j \leq 1$ for each edge $e = (v_i, v_j) \in E$. Furthermore we introduce an objective function in form $\sum_{x_i} x_i w(v_i)$.

It is obvious that the objective function and all the constraints consitute a MIP instance, and that solving it finds the maximum weigthed IS in $G$.

\section{Finding ID Sets With GLPK}

By chaining these two translations we can create a MIP formulation for a given set of AMs from a document. Solving this MIP instance will give us the optimal ID set for this document.

GLPK \cite{glpk} is a multi-platform, multi-purpose solver well suited for this task. It uses the Simplex method to solve LP problems and \textit{Branch \& Bound} (see \cite{land60a} for details) for MIP. An advantage of using \textit{Branch \& Bound} is that while traversing the \textit{search tree} it finds intermittent, sub-optimal solutions. It is thus possible to limit the total search time and instead of the optimum take the best solution found so far.\\

We will now demonstrate the full process of finding the optimal ID set of an example XML file using GLPK.

\subsection{Example}

Consider again our XML file fragment.

@@

Recall that attribute mappings in this example are $C = \{M_{y}^{a}, M_{y}^{b}, M_{y}^{c}, M_{z}^{a}\}$. Corresponding vertices in the IS formulation will be $V = \{v_{y-a}, v_{y-b}, v_{y-c}, v_{z-a}\}$. Edges in the IS formulation will be the following.

\begin{eqnarray*}
(v_{y-a},v_{y-b}) \\
(v_{y-a},v_{y-c}) \\
(v_{y-b},v_{y-c}) \\
(v_{y-a},v_{z-a}) \\
\end{eqnarray*}

First three edges are due to the type collision ($y$), the last one is due to $\iota(M_{y}^{a}) \cap \iota(M_{z}^{a}) = \{1\}$. The graph $G$ constructed in this way is in Figure \ref{image-mip-is-graph}.

\begin{figure}
  \caption{IS Representation Graph}
  \label{image-mip-is-graph}
  \centering
	\includegraphics[width=.25\textwidth]{images/is-representation}
\end{figure}

The next step is the MIP formulation. We do not need to translate from the IS formulation, as the translation from ID set formulation is straightforward, too. For each AM $m$ there will be one binary variable $x_{name(m)}$. Objective function coefficients in vector $\mathbf{c}$ will be weights of respective mappings. For each pair of AMs $m_1, m_2$ that cannot share the same ID set there shall be a row in matrix $A$ representing the inequality $x_{name(m_1)} + x_{name(m_2)} \leqslant 1$. $\mathbf{b}$ will be a vector of ones with corresponding length.

@@

The problem now is, recall, to solve the following.

@@

In GLPK \textit{MathProg} language, this translates to the following formulation. % TODO cite mathprog

\begin{scriptsize}
\begin{verbatim}
set AMs;
param Weight {i in AMs};
var x {i in AMs} binary;
maximize z: sum {i in AMs} x[i] * Weight[i];
s.t. c1: x['y-a'] + x['y-b'] <= 1;
s.t. c2: x['y-a'] + x['y-c'] <= 1;
s.t. c3: x['y-b'] + x['y-c'] <= 1;
s.t. c4: x['y-a'] + x['z-a'] <= 1;
data;
set AMs := y-a y-b y-c z-a;
param Weight :=
y-a 0.6
y-b 0.2
y-c 0.2
z-a 0.4;
end;
\end{verbatim}
\end{scriptsize}

@@

This trivial heuristics sorts all candidate AMs by their decreasing weights (Section \ref{section-definitions-weight}) and then tries to iteratively add them to solution, if possible. This way it can create only one solution, and again, this is a hungry heuristic.
